### 2014-11-19 question_sheet/questions.tex
Changes to questions.tex file

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{fancyhdr}
\usepackage[margin=2.5cm]{geometry}
\setlength{\parindent}{0in}

\begin{document}

\begin{center}
{\bf\Large {Pattern Avoidance}}
\

\

{\bf \large {Yajit Jain, Deepak Narayanan, Leon Zhang}}
\end{center}

\

\

\emph{Question 1}

\

Give two permutations in $S_3$ that 51234 avoids.

\vspace{4cm}

\emph{Question 2}

\

What is the flip of $163542$? For what $\pi \in S_6$ is the sequence $(a_n(\pi))$ definitely equal to $(a_n(163542))$?

\vspace{4cm}

\emph{Question 3}

\

Give two permutations of $S_6$ that are in $T_6$, and one permutation that is not.
\end{document}