### 2014-12-05 project3/paper/paper.tex
Changes to paper

\documentclass[11pt,letterpaper,twoside,english]{article}

\usepackage[margin=1.4in]{geometry} % controls the size of the margins

% Special symbols, etc.
\usepackage{amssymb,amsbsy,latexsym,ytableau}
\usepackage{amsmath}
\usepackage{graphics, subfigure, float} 
\usepackage{cancel}
%\usepackage{todonotes}
% Encoding settings
\usepackage[latin1]{inputenc}
\usepackage[american]{babel}
\usepackage[T1]{fontenc} 
\usepackage{tikz}

\usepackage{titling} % allows posttitle command

% AMS Math packages

\usepackage{amscd,amsthm}

\usepackage{verbatim, comment} % can comment out text 
\usepackage{mdwlist}

% Graphics
%\usepackage[dvips]{graphicx,epsfig,color}
%\usepackage{subfigure}
%\usepackage{pst-all}
%\usepackage{pstricks-add}
\usepackage{hyperref}  % can only be used with pdflatex - gives hyperlinks
\usepackage{bm} % bold math font
\usepackage{bbm}
\usepackage{pgfplots}

\usepackage{todonotes}

\newtheoremstyle{theorem}{1em}{1em}{\slshape}{0pt}{\bfseries}{.}{ }{}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{claim*}{Claim}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{remark*}{Remark}
\newtheorem{algorithm}{Algorithm}
\newtheorem*{question*}{Question}
\newtheorem{question}{Question}
\newtheorem{example}[theorem]{Example}

\providecommand{\R}{\mathbb{R}}

\providecommand{\setN}{\mathbb{N}}
\providecommand{\setZ}{\mathbb{Z}}
\providecommand{\setQ}{\mathbb{Q}}
\providecommand{\setR}{\mathbb{R}}
\providecommand{\E}{\mathrm{E}}
\providecommand{\Pr}{\mathrm{Pr}}
\providecommand{\Var}{\mathrm{Var}}

\tikzset
{
    treenode/.style = {circle, draw=black, align=center, minimum size=1cm},
}

\makeatother

\title{Almost orthogonal vectors}

\author{Yajit Jain, Deepak Narayanan, Leon Zhang}

\begin{document}

\maketitle

\section{Introduction}
Consider a collection of $N$ unit vectors $v_1, \ldots, v_N$ in $\mathbb R^n$. Define $\epsilon$ as
\[\epsilon=\max_{i\neq j}|v_i\cdot v_j|^2.\]
In this paper, we consider the problem of minimizing $\epsilon$ for various $n$ and $N$.

For $n=N$, it is clear that the minimum value of $\epsilon$ is zero, obtained by letting $v_1,\ldots, v_N$ be any orthonormal basis. For fixed $n$, it is also easy to see that the minimum value of $\epsilon$ is nondecreasing as $N$ increases: that is, if $\epsilon_1$ is the minimum value of $\epsilon$ for $N$ unit vectors in $\mathbb R^n$, and $\epsilon_2$ is the minimum value of $\epsilon$ for $N+1$ unit vectors in $\mathbb R^n$, we must have $\epsilon_1\leq \epsilon_2$. Otherwise, given a configuration of $N+1$ vectors in $\mathbb R^n$ with $\epsilon=\epsilon_2$, we could simply remove one vector and obtain a configuration of $N$ vectors whose $\epsilon$ would be less than or equal to $\epsilon_1$.

Furthermore, given a configuration of $N$ vectors in $n$-space with corresponding $\epsilon$, we can obtain a collection of $N+1$ vectors in $\mathbb R^{n+1}$ with the same $\epsilon$: simply embed the first $N$ vectors in $\mathbb R^{n}$, then add any vector perpendicular to the hyperplane in which the $N$ vectors lie. Hence the minimum $\epsilon$ for collections of $N+1$ vectors in $\mathbb R^{n+1}$ is upper-bounded by the corresponding $\epsilon$ for $N$ vectors in $\mathbb R^n$.

In this paper, we shall see that, for $n=2$, the minimum value of $\epsilon$ will be given by 
\[\epsilon=\cos^2\pi/N.\]

We shall also see that, for $N=n+1$, the minimum value of $\epsilon$ will be upper bounded by
\[\frac{1}{n^2},\] 
and this configuration is produced when the $n+1$ vectors point to the vertices of the $n$-simplex.

\section{$n=2$, generic $N$}
In this section, we consider the problem of almost orthogonal vectors of dimensionality $2$.

It is obvious that for $N=2$, we can find an $\epsilon$ equal to $0$, since it easy to pick two unit vectors $v_1$ and $v_2$ of unit length that are orthogonal to each other -- a simple example of such vectors is $[1, 0]^T$ and $[0, 1]^T$.

However, we see that the problem gets harder for larger $N$. Let us first consider the specific case of $N=3$; once we build some intuition for the problem, we will try to generalize to generic $N$.
\\

It's clear that when $N=3$, the value of $\epsilon$ must be greater than $0$, since there is no way three dimensionality-$2$ vectors can be orthogonal to each other.

In addition, we see that for $N=3$, the quantity $\epsilon$ for three unit-length vectors $v_1, v_2, v_3 \in \mathbb{R}^2$ is given by $$\max \{ |v_1 \cdot v_2|^2, |v_1 \cdot v_3|^2, |v_2 \cdot v_3|^2 \}$$

Since we don't care about the sign of the dot product between any two vectors, if we define the unit vectors $v_4, v_5, v_6 \in \mathbb{R}^2$ to be $-v_1, -v_2, -v_3$ respectively, then we see that $\epsilon$ can be equivalently expressed as $$\max_{i \neq j} \{|v_i \cdot v_j |^2 \}$$
where $i, j \in \{1,2,\ldots,6\}$ -- the above result holds because $$|v \cdot v'| = |(-v) \cdot v'| = |(-v) \cdot (-v')| = |v \cdot (-v')|$$ for some arbitrary vector $v'$. Furthermore, we see that there now exists an axis of symmetry -- one of the diagonals passing through either $v_1$ and $v_4 = -v_1$ or $v_2$ and $v_5 = -v_2$ or $v_3$ and $v_6 = -v_3$ -- so we don't need to look at the entire unit circle anymore, and instead can focus on the four vectors contained in a semi-circle resting on one of these three diameters.

Without loss of generality, let $x_1 = [1, 0]^T$; then since $x_4$ is $-x_1$, we see that $x_4 = [-1, 0]^T$. Also, without loss of generality let us assume that $x_2$ and $x_3$ are above the $x$-axis, and that $x_2$ is to the right of $x_3$.

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (1.5, -2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (1.5, -2.59) -- (0, 0);
        \draw [thick,-,blue] (-1.5, -2.59) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (2.8, -3.1) {\text{$x_2 = [\cos \pi/3,\sin \pi/3]$}};
        \node at (-2.8, -3.1) {\text{$x_3 = [\cos 2\pi/3,\sin 2\pi/3]$}};
    \end{tikzpicture}
    \caption{A configuration of three unit vectors that produce the optimum $\epsilon$ for $n=2, N=3$}
\end{figure}

Let $\theta_1$ be the angle between $x_1$ and $x_2$, $\theta_2$ be the angle between $x_2$ and $x_3$ and $\theta_3$ be the angle between $x_3$ and $x_4$. Now, we see that $\theta_1 + \theta_2 + \theta_3 = \pi$ (since vectors $v_1$ and $v_4 = -v_1$ are anti-parallel) and that $$\epsilon = \max_{i \in \{1,2,3\}} \cos ^2 \theta_i$$, since $|x_i \cdot x_j| = \cos \theta$ where $\theta$ is the angle between the two vectors $x_i$ and $x_j$.

Before proceeding, we state and prove the following lemma.
\begin{lemma}
Consider $n$ angles $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi]$ s.t. $\theta_1 + \theta_2 + \ldots + \theta_n = \pi$. Then $\epsilon = \max_i \cos^2 \theta_i$ must equal $\cos^2 \theta_j$ where $j = \text{argmin }\theta_i$.
\end{lemma}

\begin{proof}

\begin{figure}[!h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[%
			axis x line=center, axis y line=center,
			width=10cm,
			height=4cm,
			scale only axis,
			xmin=-5,
			xmax=5,
			xtick={1.57,  3.14, 4.71},
			xticklabels={, $\pi$, },
			extra x ticks={-4.71, -3.14, -1.57},
			extra x tick labels={, $-\pi$,},
			extra x tick style={
			    xticklabel style={yshift=0.5ex, anchor=south}
			},
			ymin=-1.4,
			ymax=1.4,
			ytick={-1,  0,  1}]]
			\addplot[domain=-2*pi:2*pi,smooth] (\x,{cos(\x r)});
		\end{axis}
	\end{tikzpicture}
	\caption{Plot of $\cos \theta$ versus $\theta$}
\end{figure}

Our arguments will hinge on the fact that for all $\theta \in [0, \pi/2]$, the function $\cos \theta$ is decreasing -- this is easy to see from Figure $2$. Without loss of generality, let us assume that $\theta_1 \geq \theta_2 \geq \ldots \geq\theta_n$. Hence our theorem statement is equivalent to proving that $\epsilon$ is equal to $\cos^2 \theta_n$.

We split our proof into two cases,
\begin{itemize}
\item $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi/2]$: In this case, it is easy to see that $\epsilon = \cos^2 \theta_n$ from the fact that $\cos \theta$ is a decreasing function in $\theta$ if $\theta \in [0, \pi/2]$.

\item One of $\theta_1, \theta_2, \dots, \theta_n$ is greater than $\pi/2$:

Then, if $\theta_1 > \pi/2$ we see that $\cos^2 \theta_1 = \cos^2 (\pi - \theta_1)$ is equal to $\cos^2 (\theta_2 + \theta_3 + \ldots + \theta_n)$. Furthermore, $\pi - \theta_1 = \theta_2 + \theta_3 + \ldots + \theta_n < \pi/2$, which means $\cos^2 \theta_1 = \cos^2 (\theta_2 + \theta_3 + \ldots + \theta_n) < \cos^2 \theta_n$. (since $\theta_n < \theta_2 + \theta_3 + \ldots + \theta_n$)

In addition, we see that for all $i \in \{2, 3, \ldots, n-1\}$, $\cos^2 \theta_i < \cos^2 \theta_n$, from which we can conclude that in this case too, $\epsilon$ is equal to $\cos^2 \theta_n$.

\end{itemize}
\end{proof}

From the above lemma, we can conclude that if $j = \text{argmin } \theta_i$ and $\theta_1 + \theta_2 + \theta_3 = \pi$, then $\epsilon = \cos^2 \theta_j$.

Furthermore, we can obtain an upper bound on $\theta_j$ by observing that $\theta_ 1 + \theta_ 2 + \theta_3 \geq \theta_ j + \theta_j + \theta_j = 3 \theta_j \Rightarrow \theta_j \leq \pi/3$. Since we're interested in the smallest such $\epsilon$ and since the cosine function is a decreasing function in $\theta$ between $0$ and $\pi/2$, we conclude that the optimum value of $\epsilon$ for $n=2$ and $N=3$ is $\cos^2 \pi / 3$. Figure $1$ shows this optimum configuration.

Given the above result for $N=3$, we attempt to generalize to any integer $N$ in the following theorem.

\begin{theorem}
For $n=2$ and arbitrary $N$, the optimum value of $\epsilon$ is given by $\cos^2 \pi/N$.
\end{theorem}

\begin{proof}
The proof of this theorem is similar to the proof for the specific case of $N=3$.

Again without loss of generality, we can assume that the vectors $x_1, x_2, \ldots, x_N$ are on or above the $x$-axis, and that $x_1 = [1,0]^T$ and that $x_{N+1} = [-1,0]^T$ -- if any vector $x_i$ were not above the $x$-axis, then we could just consider $-x_i$ instead.

Let us define $\theta_i$ as the angle between the vectors $x_i$ and $x_{i+1}$. Then, we quickly see that $\theta_1 + \theta_2 + \ldots + \theta_N = \pi$. If $j = \text{argmin } \theta_i$, $\epsilon = \cos^2 \theta_j$ by the above Lemma.

Since $\theta_ 1 + \theta_2 + \ldots + \theta_N \geq N \cdot \theta_j \Rightarrow \theta_j \leq \pi/N$, we conclude that the optimum $\epsilon$ value is in fact equal to $\cos^2 \pi/N$ (again making use of the fact that the cosine function is decreasing).
\end{proof}

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (-0.93, -2.85)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, -2.85) -- (0, 0);
        \draw [thick,-,blue] (-2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (-0.93, -2.85) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (4.5, -1.76) {\text{$x_2 = [\cos \pi/5,\sin \pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (-3.43, -3) {\text{$x_4 = [\cos 3\pi/5,\sin 3\pi/5]$}};
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$}
\end{figure}

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, 1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, 2.85)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-2.42, 1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, -2.85) -- (0, 0);
        \draw [thick,-,blue] (-2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, 2.85) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (-4.5, 1.76) {\text{$x'_2 = [\cos 6\pi/5,\sin 6\pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (3.43, 3) {\text{$x'_4 = [\cos 8\pi/5,\sin 8\pi/5]$}};

        \draw [thick,-,red,dashed] (3, 0) -- (0.93, 2.85);
        \draw [thick,-,red,dashed] (3, 0) -- (0.93, -2.85);
        \draw [thick,-,red,dashed] (0.93, 2.85) -- (-2.42, 1.76);
        \draw [thick,-,red,dashed] (0.93, -2.85) -- (-2.42, -1.76);
        \draw [thick,-,red,dashed] (-2.42, 1.76) -- (-2.42, -1.76);
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$ with $x_2$ and $x_4$ from the previous figure reversed}
\end{figure}

Figure $3$ shows an optimum configuration for $n=2$ and $N=5$. Note by taking reflections of $x_2$ and $x_4$ in Figure $3$, we get five vectors that form a regular $5$-gon -- this can be seen in Figure $4$. A similar transformation can be done for all odd $N$; unfortunately, even $N$ cannot produce regular polygons.

\section{The Parallelogram Law}

First we give the definition of the norm on vectors in $\R^n$ in terms of the dot product:

\begin{definition}
For a vector $v\in\R^n$ the norm of $v$ is the positive value of $||v||$ given by the equation
$$
||v||^2=v\cdot v.
$$
\end{definition}
\begin{theorem}[Parallelogram Law]
For vectors $v_1,\cdots,v_n\in\R^n$, the following identity holds
$$
||v_1+\cdots+v_n||^2=\displaystyle\sum_{i=1}^n||v_i||^2+2\displaystyle\sum_{1\le i<j\le n}v_i\cdot v_j.
$$
\proof
First consider the following identity from the definition of the norm for vectors $u,v\in\R^n$:
$$
||u+v||^2=(u+v)\cdot (u+v)=u\cdot u+u\cdot v+ v\cdot u + v\cdot v=||u||^2+||v||^2+2u\cdot v.
$$
Now if we let $u=v_1$ and $v= v_2+\cdots+v_n$ we get 
$$
||v_1+v_2+\cdots+v_n||^2=||v_1||^2+||v_2+\cdots +v_n||^2+2\sum_{i=2}^n v_1\cdot v_i.
$$
If we recurse on $||v_2+\cdots +v_n||^2$ using induction we get the desired result. 
\end{theorem}

\section{$N = n+1$}
Let us consider the situation in which we try to minimize $\epsilon$ for $n+1$ vectors in an $n$ dimensional space. First we introduce the definition of a simplex, and then we will proceed to characterize $\epsilon$.

\begin{definition}
An $n$-dimensional simplex is given by the convex hull of the $n+1$ points in $\R^{n+1}$ described as having a 1 in a single coordinate and zeros in every other coordinate. 
\end{definition}

Figure $5$ shows diagrams of a 2-dimensional simplex.

\begin{figure}[!h]
    \centering
        \begin{tikzpicture}[yscale=-1, scale=0.75]
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % z-axis
        \draw [thick, ->] (3.0, -2.0) -- (-3.0, 2.0);
        % origin label
        \node at (-1, -0.3) {\text{$(0, 0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
         % z-axis label
        \node at (-3.4, 1.9) {\text{$z$}};
        \node at (3.5, -0.4) {\text{$(1, 0, 0)$}};
        \node at (1, -3.4) {\text{$(0, 1, 0)$}};
        \node at (-2.8, 1.1) {\text{$(0, 0, 1)$}};
        % circle
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (0, -3)[blue,fill=blue] circle (0.1cm);
        \draw (-1.8, 1.2)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,red,dashed] (3, 0) -- (0, -3);
        \draw [thick,-,red,dashed] (0, -3) -- (-1.8, 1.2);
        \draw [thick,-,red,dashed] (3, 0) -- (-1.8, 1.2);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (0, -3) -- (0, 0);
        \draw [thick,-,blue] (-1.8, 1.2) -- (0, 0);
    \end{tikzpicture}
    \begin{tikzpicture}[yscale=-1, scale=0.75]
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.6, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, 2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,red,dashed] (3, 0) -- (-1.5, 2.59);
        \draw [thick,-,red,dashed] (3, 0) -- (-1.5, -2.59);
        \draw [thick,-,red,dashed] (-1.5, -2.59) -- (-1.5, 2.59);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-1.5, 2.59) -- (0, 0);
        \draw [thick,-,blue] (-1.5, -2.59) -- (0, 0);
    \end{tikzpicture}

    \caption{A 2-dimensional simplex drawn in $\R^3$ on the left, and then centered at the origin in $\R^2$ on the right.}
\end{figure}

Before proceeding with the theorem, we notice that when $n=2$, the configuration of $N=3$ vectors that produced the minimum value of $\epsilon$ made up a simplex centered at the origin.

\begin{theorem}
If $N=n+1$ and the vectors in question are denoted $v_1,\cdots, v_N$, then over all configurations of these vectors the maximum value of $v_i\cdot v_j$ for $i\neq j$ is $\frac{-1}{n}$. This is achieved when the vectors are arranged in an $n$ dimensional simplex centered at the origin, and in this configuration $\epsilon=\frac{1}{n^2}$.  
\end{theorem}

\proof
Let $\delta= \max_{i\neq j}\{v_i\cdot v_j\}$ for a given set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, and let $\delta_{min}$ be the minimum possible $\delta$ obtained across all unit vectors $v_1, v_2, \ldots, v_{n+1}$. We first show that $\delta_{min} \le-\frac{1}{n}$ by showing that the origin-centered simplex configuration of the vectors $v_1,\cdots v_{n+1}$ produces a value $\delta=-\frac{1}{n}$, and then will later show that $\delta_{min} \geq -\frac{1}{n}$ as well.

We must first describe the vectors that make up the origin centered simplex. To center our simplex at the origin we take each vector in the original simplex, and subtract from it the vector representing the centroid of the original simplex. The centroid of the original simplex takes the form $(\frac{1}{N},\frac{1}{N},\ldots,\frac{1}{N})$, so the points of the new simplex take the form
$$
\left(-\frac{1}{N},\cdots,-\frac{1}{N},\frac{N-1}{N},-\frac{1}{N},\cdots,-\frac{1}{N}\right).
$$
If we renormalize these vectors so that each vector has unit length, we get 
$$
\sqrt{\frac{N}{N-1}}\left(-\frac{1}{N},\cdots,-\frac{1}{N},\frac{N-1}{N},-\frac{1}{N},\cdots,-\frac{1}{N}\right).
$$
Now, taking the dot product between any two such vectors gives us
$$
\frac{N}{N-1}\left(-2\cdot\frac{N-1}{N^2}+(N-2)\cdot\frac{1}{N^2}\right)=\frac{N}{N-1}\cdot-\frac{1}{N}=-\frac{1}{N-1}=-\frac{1}{n}.
$$
So an upper bound on the value of $\delta_{min}$ for all configurations of $n+1$ vectors in $n+1$ dimensional space is $\frac{-1}{n}$.

To see that $\delta_{min} \ge-\frac{1}{n}$ we use the parallelogram law.

Recall that the parallelogram law states that

$$
||v_1+\cdots+v_{n+1}||^2=\displaystyle\sum_{i=1}^{n+1}||v_i||^2+2\displaystyle\sum_{1\le i<j\le n+1}v_i\cdot v_j.
$$
In our case we know three things. First, since the norm of any vector is greater than or equal to $0$, we can conclude that $||v_1 + \cdots + v_{n+1}||^2 \geq 0$. Second, $v_i\cdot v_j\le\delta$ for any $i<j$. Third, $||v_i||=1$. Therefore 
$$
0\le n+1+n(n+1)\delta.
$$
This implies that $\delta\ge -\frac{1}{n}$ for any set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, from which we can conclude that $\delta_{min} \ge -\frac{1}{n}$ as well.

Since $\delta_{min} \le -\frac{1}{n}$ and $\delta_{min} \ge -\frac{1}{n}$, we can conclude that $\delta_{min}=-\frac{1}{n}$. From this, we can see that a $\epsilon = \frac{1}{n^2}$ is achievable for the general case of when $N=n+1$ -- this leads us to conclude that the minimum possible $\epsilon$ value for $N=n+1$ is at most $\frac{1}{n^2}$.
\section{$N>n+1$}

We conjecture that in cases where $N>n+1$ we can never have an arrangement of the vectors where all values of $|v_i\cdot v_j|^2$ are the same. Proving this remains an object of future study.

\section{Applications to sphere packing}
We should be able to prove something about the radius of spheres you can pack in the case of $N=n+1$, taking as intuition square and hexagonal tilings of the plane and cubic and tetrahedral tilings of 3-space.

\end{document}