### 2014-12-09 project3/paper/paper.tex
Changes to the paper -- added code for algorithm, and also corrected error in proof.

\documentclass[11pt,letterpaper,twoside,english]{article}

\usepackage[margin=1.4in]{geometry} % controls the size of the margins

% Special symbols, etc.
\usepackage{amssymb,amsbsy,latexsym,ytableau}
\usepackage{amsmath}
\usepackage{graphics, subfigure, float} 
\usepackage{cancel}
%\usepackage{todonotes}
% Encoding settings
\usepackage[latin1]{inputenc}
\usepackage[american]{babel}
\usepackage[T1]{fontenc} 
\usepackage{tikz}
\usepackage[scaled]{beramono}
\usepackage{listings}

\lstset{
  language=Python,
  showstringspaces=false,
  formfeed=\newpage,
  tabsize=4,
  commentstyle=\itshape,
  basicstyle=\ttfamily,
  morekeywords={while, if}
  captionpos=b,
}

\usepackage{titling} % allows posttitle command

% AMS Math packages

\usepackage{amscd,amsthm}

\usepackage{verbatim, comment} % can comment out text 
\usepackage{mdwlist}

% Graphics
%\usepackage[dvips]{graphicx,epsfig,color}
%\usepackage{subfigure}
%\usepackage{pst-all}
%\usepackage{pstricks-add}
\usepackage{hyperref}  % can only be used with pdflatex - gives hyperlinks
\usepackage{bm} % bold math font
\usepackage{bbm}
\usepackage{pgfplots}

\renewcommand{\lstlistingname}{Algorithm}

\usepackage{todonotes}
\usepackage{caption}
\DeclareCaptionFormat{myformat}{#1#2#3\hrulefill}
\captionsetup{format=myformat}
\captionsetup[lstlisting]{position=bottom,format=myformat}

\newtheoremstyle{theorem}{1em}{1em}{\slshape}{0pt}{\bfseries}{.}{ }{}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{claim*}{Claim}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{remark*}{Remark}
\newtheorem{algorithm}{Algorithm}
\newtheorem*{question*}{Question}
\newtheorem{question}{Question}
\newtheorem{example}[theorem]{Example}

\providecommand{\R}{\mathbb{R}}

\providecommand{\setN}{\mathbb{N}}
\providecommand{\setZ}{\mathbb{Z}}
\providecommand{\setQ}{\mathbb{Q}}
\providecommand{\setR}{\mathbb{R}}
\providecommand{\E}{\mathrm{E}}
\providecommand{\Pr}{\mathrm{Pr}}
\providecommand{\Var}{\mathrm{Var}}

\tikzset
{
    treenode/.style = {circle, draw=black, align=center, minimum size=1cm},
}

\makeatother

\title{Almost orthogonal vectors}

\author{Yajit Jain, Deepak Narayanan, Leon Zhang}

\begin{document}

\maketitle

\section{Introduction}
Consider a collection of $N$ unit vectors $v_1, \ldots, v_N$ in $\mathbb R^n$. Define $\epsilon$ as
\[\epsilon=\max_{i\neq j}|v_i\cdot v_j|^2.\]
In this paper, we consider the problem of minimizing $\epsilon$ for various $n$ and $N$.

For $n=N$, it is clear that the minimum value of $\epsilon$ is zero, obtained by letting $v_1,\ldots, v_N$ be any orthonormal basis. For fixed $n$, it is also easy to see that the minimum value of $\epsilon$ is nondecreasing as $N$ increases: that is, if $\epsilon_1$ is the minimum value of $\epsilon$ for $N$ unit vectors in $\mathbb R^n$, and $\epsilon_2$ is the minimum value of $\epsilon$ for $N+1$ unit vectors in $\mathbb R^n$, we must have $\epsilon_1\leq \epsilon_2$. Otherwise, given a configuration of $N+1$ vectors in $\mathbb R^n$ with $\epsilon=\epsilon_2$, we could simply remove one vector and obtain a configuration of $N$ vectors whose $\epsilon$ would be less than or equal to $\epsilon_1$.

Furthermore, given a configuration of $N$ vectors in $n$-space with corresponding $\epsilon$, we can obtain a collection of $N+1$ vectors in $\mathbb R^{n+1}$ with the same $\epsilon$: simply embed the first $N$ vectors in $\mathbb R^{n}$, then add any vector perpendicular to the hyperplane in which the $N$ vectors lie. Hence the minimum $\epsilon$ for collections of $N+1$ vectors in $\mathbb R^{n+1}$ is upper-bounded by the corresponding $\epsilon$ for $N$ vectors in $\mathbb R^n$.

In this paper, we shall see that, for $n=2$, the minimum value of $\epsilon$ will be given by 
\[\epsilon=\cos^2\pi/N.\]

We shall also see that, for $N=n+1$, the minimum value of $\epsilon$ is upper bounded by
\[\frac{1}{n^2},\] 
and this configuration is produced when the $n+1$ vectors point to the vertices of the $n$-simplex.

\section{$n=2$, generic $N$}
In this section, we consider the problem of almost orthogonal vectors of dimensionality $2$.

It is obvious that for $N=2$, we can find an $\epsilon$ equal to $0$, since it easy to pick two unit vectors $v_1$ and $v_2$ of unit length that are orthogonal to each other -- a simple example of such vectors is $[1, 0]^T$ and $[0, 1]^T$.

However, we see that the problem gets harder for larger $N$. Let us first consider the specific case of $N=3$; once we build some intuition for the problem, we will try to generalize to generic $N$.
\\

It's clear that when $N=3$, the value of $\epsilon$ must be greater than $0$, since there is no way three dimensionality-$2$ vectors can be orthogonal to each other. We make this more concrete in the following theorem.

\begin{theorem}
For $n=2$ and $N=3$, the minimum possible $\epsilon$ is equal to $1/4$.
\end{theorem}

Before proving the above theorem, we state and prove the following lemma.

\begin{lemma}
Consider $n$ angles $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi]$ s.t. $\theta_1 + \theta_2 + \ldots + \theta_n = \pi$. Then $\alpha = \max_i \cos^2 \theta_i$ must equal $\cos^2 \theta_j$ where $j = \text{argmin }\theta_i$.
\end{lemma}

\begin{proof}

\begin{figure}[!h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[%
			axis x line=center, axis y line=center,
			width=10cm,
			height=4cm,
			scale only axis,
			xmin=-5,
			xmax=5,
			xtick={1.57,  3.14, 4.71},
			xticklabels={, $\pi$, },
			extra x ticks={-4.71, -3.14, -1.57},
			extra x tick labels={, $-\pi$,},
			extra x tick style={
			    xticklabel style={yshift=0.5ex, anchor=south}
			},
			ymin=-1.4,
			ymax=1.4,
			ytick={-1,  0,  1}]]
			\addplot[domain=-2*pi:2*pi,smooth] (\x,{cos(\x r)});
		\end{axis}
	\end{tikzpicture}
	\caption{Plot of $\cos \theta$ versus $\theta$}
\end{figure}

Our arguments will hinge on the fact that for all $\theta \in [0, \pi/2]$, the function $\cos \theta$ is decreasing -- this is easy to see from Figure $1$. Without loss of generality, let us assume that $\theta_1 \geq \theta_2 \geq \ldots \geq\theta_n$. Hence our theorem statement is equivalent to proving that $\alpha$ is equal to $\cos^2 \theta_n$.

We split our proof into two cases,
\begin{itemize}
\item $\theta_1, \theta_2, \ldots, \theta_n \in [0, \pi/2]$: In this case, it is easy to see that $\alpha = \cos^2 \theta_n$ from the fact that $\cos \theta$ is a decreasing function in $\theta$ if $\theta \in [0, \pi/2]$.

\item One of $\theta_1, \theta_2, \dots, \theta_n$ is greater than $\pi/2$:

Then, if $\theta_1 > \pi/2$ we see that $\cos^2 \theta_1 = \cos^2 (\pi - \theta_1)$ is equal to $\cos^2 (\theta_2 + \theta_3 + \ldots + \theta_n)$. Furthermore, $\pi - \theta_1 = \theta_2 + \theta_3 + \ldots + \theta_n < \pi/2$, which means $\cos^2 \theta_1 = \cos^2 (\theta_2 + \theta_3 + \ldots + \theta_n) < \cos^2 \theta_n$. (since $\theta_n < \theta_2 + \theta_3 + \ldots + \theta_n$)

In addition, we see that for all $i \in \{2, 3, \ldots, n-1\}$, $\cos^2 \theta_i < \cos^2 \theta_n$, from which we can conclude that in this case too, $\alpha$ is equal to $\cos^2 \theta_n$.

\end{itemize}
\end{proof}

Given this lemma, we now prove the above theorem.

\begin{proof}
Observe that for $N=3$, the quantity $\epsilon$ for three unit-length vectors $v_1, v_2, v_3 \in \mathbb{R}^2$ is given by $$\max \{ |v_1 \cdot v_2|^2, |v_1 \cdot v_3|^2, |v_2 \cdot v_3|^2 \}$$

Since we don't care about the sign of the dot products between any two vectors (we're looking at the squares of the dot products in the above equation), and since $$|v \cdot v'| = |v \cdot (-v')|$$ for some arbitrary vector $v'$, we see that it's easy to transform the three vectors $v_1$, $v_2$ and $v_3$ so that they all lie in the same semi-circle -- to accomplish this, at most one vector needs to be reflected about the origin.

Without loss of generality, let $v_1 = [1, 0]^T$. Also, without loss of generality let us assume that $v_2$ and $v_3$ are above the $x$-axis, and that $v_1$, $v_2$ and $v_3$ are in anti-clockwise order.

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (1.5, -2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (1.5, -2.59) -- (0, 0);
        \draw [thick,-,blue] (-1.5, -2.59) -- (0, 0);

        \node at (4, -0.3) {\text{$v_1 = [1,0]$}};
        \node at (2.8, -3.1) {\text{$v_2 = [\cos \pi/3,\sin \pi/3]$}};
        \node at (-2.8, -3.1) {\text{$v_3 = [\cos 2\pi/3,\sin 2\pi/3]$}};
    \end{tikzpicture}
    \caption{A configuration of three unit vectors that produce the optimum $\epsilon$ for $n=2, N=3$}
\end{figure}

Let $\theta_1$ be the angle between $v_1$ and $v_2$ and $\theta_2$ be the angle between $v_2$ and $v_3$. Let us now define $\theta_3$ such that $\theta_1 + \theta_2 + \theta_3 = \pi$. Observe that $|v_1 \cdot v_3|^2 = \cos^2(\theta_1 + \theta_2) = \cos^2 (\pi - \theta_3) = \cos^2 \theta_3$, hence we can conclude that

$$\epsilon = \max_{i \in \{1,2,3\}} \cos ^2 \theta_i$$

From the above lemma, we can conclude that if $j = \text{argmin } \theta_i$ and $\theta_1 + \theta_2 + \theta_3 = \pi$, then $\epsilon = \cos^2 \theta_j$.

Furthermore, we can obtain an upper bound on $\theta_j$ by observing that $\theta_ 1 + \theta_ 2 + \theta_3 \geq \theta_ j + \theta_j + \theta_j = 3 \theta_j \Rightarrow \theta_j \leq \pi/3$. Since we're interested in the smallest such $\epsilon$ and since the cosine function is a decreasing function in $\theta$ between $0$ and $\pi/2$, we conclude that the optimum value of $\epsilon$ for $n=2$ and $N=3$ is $\cos^2 \pi / 3$. Figure $1$ shows this optimum configuration.

\end{proof}

Given the above result for $N=3$, we attempt to generalize to any integer $N$ in the following theorem.

\begin{theorem}
For $n=2$ and arbitrary $N$, the optimum value of $\epsilon$ is given by $\cos^2 \pi/N$.
\end{theorem}

\begin{proof}
The proof of this theorem is similar to the proof for the specific case of $N=3$.

Again without loss of generality, we can assume that the vectors $v_1, v_2, \ldots, v_N$ are on or above the $x$-axis, and that $v_1 = [1,0]^T$ -- if any vector $v_i$ were not above the $x$-axis, then we could just consider $-v_i$ (the reflection of $v_i$ about the $x$-axis) instead.

Let us define $\theta_i$ as the angle between the vectors $v_i$ and $v_{i+1}$ for $i \in \{1, 2, \ldots, N-1\}$, and let $\theta_N$ be the angle such that $\theta_1 + \theta_2 + \ldots + \theta_N = \pi$.

Observe that as before, we are interested in maximizing the square of the cosine of the angle between any two vectors in $v_1, v_2, \ldots, v_N$. Note that here, the angle between any two vectors in $v_j$ and $v_k$ in $v_1, v_2, \ldots, v_N$ can be expressed as $\cos^2 (\sum_{i=j} ^{k-1} \theta_i)$. Note that, however, if $\sum_{i=j}^{k-1} \theta_i \leq \pi/2$, then $\cos^2 \theta_j \geq \cos^2(\sum_{i=j}^{k-1} \theta_i)$ and that if $\sum_{i=j}^{k-1} \theta_i > \pi/2$, then $\cos^2 (\sum_{i=j}^{k-1} \theta_i) = \cos^2 (\pi - \sum_{i=j}^{k-1} \theta_i)$ which is less than $\cos^2 \theta_i$ for any $i$ not in $\{j, j+1, \ldots, k-1\}$.

From this we can conclude that $\epsilon$ is equal to $\max_i \cos^2 \theta_i$. Then, if $j = \text{argmin } \theta_i$, then we see that $\epsilon = \cos^2 \theta_j$ by the above Lemma.

Since $\theta_ 1 + \theta_2 + \ldots + \theta_N \geq N \cdot \theta_j \Rightarrow \theta_j \leq \pi/N$, we conclude that the optimum $\epsilon$ value is in fact equal to $\cos^2 \pi/N$ (again making use of the fact that the cosine function is decreasing between $0$ and $\pi/2$).
\end{proof}

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (-0.93, -2.85)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, -2.85) -- (0, 0);
        \draw [thick,-,blue] (-2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (-0.93, -2.85) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (4.5, -1.76) {\text{$x_2 = [\cos \pi/5,\sin \pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (-3.43, -3) {\text{$x_4 = [\cos 3\pi/5,\sin 3\pi/5]$}};
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$}
\end{figure}

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}[yscale=-1] 
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.5, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, 1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, -2.85)[blue,fill=blue] circle (0.1cm);
        \draw (-2.42, -1.76)[blue,fill=blue] circle (0.1cm);
        \draw (0.93, 2.85)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-2.42, 1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, -2.85) -- (0, 0);
        \draw [thick,-,blue] (-2.42, -1.76) -- (0, 0);
        \draw [thick,-,blue] (0.93, 2.85) -- (0, 0);

        \node at (4, -0.3) {\text{$x_1 = [1,0]$}};
        \node at (-4.5, 1.76) {\text{$x'_2 = [\cos 6\pi/5,\sin 6\pi/5]$}};
        \node at (3.43, -3) {\text{$x_3 = [\cos 2\pi/5,\sin 2\pi/5]$}};
         \node at (-4.5, -1.76) {\text{$x_5 = [\cos 4\pi/5,\sin 4\pi/5]$}};
        \node at (3.43, 3) {\text{$x'_4 = [\cos 8\pi/5,\sin 8\pi/5]$}};

        \draw [thick,-,red,dashed] (3, 0) -- (0.93, 2.85);
        \draw [thick,-,red,dashed] (3, 0) -- (0.93, -2.85);
        \draw [thick,-,red,dashed] (0.93, 2.85) -- (-2.42, 1.76);
        \draw [thick,-,red,dashed] (0.93, -2.85) -- (-2.42, -1.76);
        \draw [thick,-,red,dashed] (-2.42, 1.76) -- (-2.42, -1.76);
    \end{tikzpicture}
    \caption{A configuration of five unit vectors that produce the optimum $\epsilon$ for $n=2, N=5$ with $x_2$ and $x_4$ from the previous figure reversed}
\end{figure}

Figure $3$ shows an optimum configuration for $n=2$ and $N=5$. Note by taking reflections of $x_2$ and $x_4$ in Figure $3$, we get five vectors that form a regular $5$-gon -- this can be seen in Figure $4$. A similar transformation can be done for all odd $N$; unfortunately, even $N$ cannot produce regular polygons.

\section{$N = n+1$}

We now look to calculate the optimum value of $\epsilon$ for the specific case of $N = n+1$. Before proving an upper bound on what the optimum value of $\epsilon$ can be, we look at the parallelogram law.

\subsection{The Parallelogram Law}

First we give the definition of the norm on vectors in $\R^n$ in terms of the dot product:

\begin{definition}
For a vector $v\in\R^n$ the norm of $v$ is the positive value of $||v||$ given by the equation
$$
||v||^2=v\cdot v.
$$
\end{definition}
\begin{theorem}[Parallelogram Law]
For vectors $v_1,\cdots,v_n\in\R^n$, the following identity holds
$$
||v_1+\cdots+v_n||^2=\displaystyle\sum_{i=1}^n||v_i||^2+2\displaystyle\sum_{1\le i<j\le n}v_i\cdot v_j.
$$
\proof
First consider the following identity from the definition of the norm for vectors $u,v\in\R^n$:
$$
||u+v||^2=(u+v)\cdot (u+v)=u\cdot u+u\cdot v+ v\cdot u + v\cdot v=||u||^2+||v||^2+2u\cdot v.
$$
Now if we let $u=v_1$ and $v= v_2+\cdots+v_n$ we get 
$$
||v_1+v_2+\cdots+v_n||^2=||v_1||^2+||v_2+\cdots +v_n||^2+2\sum_{i=2}^n v_1\cdot v_i.
$$
If we recurse on $||v_2+\cdots +v_n||^2$ using induction we get the desired result. 
\end{theorem}

\subsection{Upper bound on minimum possible $\epsilon$}
Let us consider the situation in which we try to minimize $\epsilon$ for $n+1$ vectors in an $n$ dimensional space. First we introduce the definition of a simplex, and then we will proceed to characterize $\epsilon$.

\begin{definition}
An $n$-dimensional simplex is given by the convex hull of the $n+1$ points in $\R^{n+1}$ described as having a 1 in a single coordinate and zeros in every other coordinate. 
\end{definition}

Figure $5$ shows diagrams of a 2-dimensional simplex.

\begin{figure}[!h]
    \centering
        \begin{tikzpicture}[yscale=-1, scale=0.75]
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % z-axis
        \draw [thick, ->] (3.0, -2.0) -- (-3.0, 2.0);
        % origin label
        \node at (-1, -0.3) {\text{$(0, 0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
         % z-axis label
        \node at (-3.4, 1.9) {\text{$z$}};
        \node at (3.5, -0.4) {\text{$(1, 0, 0)$}};
        \node at (1, -3.4) {\text{$(0, 1, 0)$}};
        \node at (-2.8, 1.1) {\text{$(0, 0, 1)$}};
        % circle
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (0, -3)[blue,fill=blue] circle (0.1cm);
        \draw (-1.8, 1.2)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,red,dashed] (3, 0) -- (0, -3);
        \draw [thick,-,red,dashed] (0, -3) -- (-1.8, 1.2);
        \draw [thick,-,red,dashed] (3, 0) -- (-1.8, 1.2);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (0, -3) -- (0, 0);
        \draw [thick,-,blue] (-1.8, 1.2) -- (0, 0);
    \end{tikzpicture}
    \begin{tikzpicture}[yscale=-1, scale=0.75]
        % x-axis
        \draw [thick,->] (-4.5, 0) -- (4.5, 0);
        % y-axis
        \draw [thick,->] (0, 4.5) -- (0, -4.5);
        % origin label
        \node at (-0.6, 0.3) {\text{$(0, 0)$}};
        % x-axis label
        \node at (4.5, 0.5) {\text{$x$}};
        % y-axis label
        \node at (0, -5) {\text{$y$}};
        % circle
        \draw (0,0) circle (3cm);
        \draw (3,0)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, 2.59)[blue,fill=blue] circle (0.1cm);
        \draw (-1.5, -2.59)[blue,fill=blue] circle (0.1cm);

        \draw [thick,-,red,dashed] (3, 0) -- (-1.5, 2.59);
        \draw [thick,-,red,dashed] (3, 0) -- (-1.5, -2.59);
        \draw [thick,-,red,dashed] (-1.5, -2.59) -- (-1.5, 2.59);

        \draw [thick,-,blue] (3, 0) -- (0, 0);
        \draw [thick,-,blue] (-1.5, 2.59) -- (0, 0);
        \draw [thick,-,blue] (-1.5, -2.59) -- (0, 0);
    \end{tikzpicture}

    \caption{A 2-dimensional simplex drawn in $\R^3$ on the left, and then centered at the origin in $\R^2$ on the right.}
\end{figure}

Before proceeding with the theorem, we notice that when $n=2$, the configuration of $N=3$ vectors that produced the minimum value of $\epsilon$ made up a simplex centered at the origin.

\begin{theorem}
If $N=n+1$ and the vectors in question are denoted $v_1,\cdots, v_N$, then over all configurations of these vectors the maximum value of $v_i\cdot v_j$ for $i\neq j$ is $\frac{-1}{n}$. This is achieved when the vectors are arranged in an $n$ dimensional simplex centered at the origin, and in this configuration $\epsilon=\frac{1}{n^2}$.  
\end{theorem}

\begin{proof}
Let $\delta= \max_{i\neq j}\{v_i\cdot v_j\}$ for a given set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, and let $\delta_{min}$ be the minimum possible $\delta$ obtained across all unit vectors $v_1, v_2, \ldots, v_{n+1}$. We first show that $\delta_{min} \le-\frac{1}{n}$ by showing that the origin-centered simplex configuration of the vectors $v_1,\cdots v_{n+1}$ produces a value $\delta=-\frac{1}{n}$, and then will later show that $\delta_{min} \geq -\frac{1}{n}$ as well.

We must first describe the vectors that make up the origin centered simplex. To center our simplex at the origin we take each vector in the original simplex, and subtract from it the vector representing the centroid of the original simplex. The centroid of the original simplex takes the form $(\frac{1}{N},\frac{1}{N},\ldots,\frac{1}{N})$, so the points of the new simplex take the form
$$
\left(-\frac{1}{N},\cdots,-\frac{1}{N},\frac{N-1}{N},-\frac{1}{N},\cdots,-\frac{1}{N}\right).
$$
If we renormalize these vectors so that each vector has unit length, we get 
$$
\sqrt{\frac{N}{N-1}}\left(-\frac{1}{N},\cdots,-\frac{1}{N},\frac{N-1}{N},-\frac{1}{N},\cdots,-\frac{1}{N}\right).
$$
Now, taking the dot product between any two such vectors gives us
$$
\frac{N}{N-1}\left(-2\cdot\frac{N-1}{N^2}+(N-2)\cdot\frac{1}{N^2}\right)=\frac{N}{N-1}\cdot-\frac{1}{N}=-\frac{1}{N-1}=-\frac{1}{n}.
$$
So an upper bound on the value of $\delta_{min}$ for all configurations of $n+1$ vectors in $n+1$ dimensional space is $\frac{-1}{n}$.

To see that $\delta_{min} \ge-\frac{1}{n}$ we use the parallelogram law.

Recall that the parallelogram law states that

$$
||v_1+\cdots+v_{n+1}||^2=\displaystyle\sum_{i=1}^{n+1}||v_i||^2+2\displaystyle\sum_{1\le i<j\le n+1}v_i\cdot v_j.
$$
In our case we know three things. First, since the norm of any vector is greater than or equal to $0$, we can conclude that $||v_1 + \cdots + v_{n+1}||^2 \geq 0$. Second, $v_i\cdot v_j\le\delta$ for any $i<j$. Third, $||v_i||=1$. Therefore 
$$
0\le n+1+n(n+1)\delta.
$$
This implies that $\delta\ge -\frac{1}{n}$ for any set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, from which we can conclude that $\delta_{min} \ge -\frac{1}{n}$ as well.

Since $\delta_{min} \le -\frac{1}{n}$ and $\delta_{min} \ge -\frac{1}{n}$, we can conclude that $\delta_{min}=-\frac{1}{n}$. From this, we can see that a $\epsilon = \frac{1}{n^2}$ is achievable for the general case of when $N=n+1$ -- this leads us to conclude that the minimum possible $\epsilon$ value for $N=n+1$ is at most $\frac{1}{n^2}$.
\end{proof}

Note that proving that the minimum possible epsilon for the case of $N = n+1$ is equal to $1/n^2$ is much harder, and we leave this for future work.

\section{Simulated annealing and associated conjectures}

Since visualizing in spaces of dimensionality greater than $3$ is extremely hard, we used an annealing approach to compute near-optimal epsilon values for different $(n, N)$ pairs. We describe this method in greater detail below.

Our general intuition tells us that perturbing a set of nearly optimal vectors slightly could give us a new set of vectors that produce even a smaller epsilon value. Given this, we start off with a set of random vectors, and try to move these vectors towards smaller epsilon values.

At every iteration, we try adding a vector whose norm becomes smaller with every passing iteration, to the already computed optimal vector.

\begin{lstlisting}
def get_min_epsilon_annealing(n, N, num_iter, start_vectors):
    min_epsilon = 1.0  # Start off with the worst possible epsilon
    current_vectors = start_vectors
    i = 0
    while (i < num_iter):
        temperature = 1.0 / float(i + 1)

        # Get perturbing vectors
        random_vectors = get_random_vectors(n, N)

        # Now perturb current_vectors
        new_vectors = list()
        for j in xrange(N):
            new_vector = list()
            for k in xrange(n):
                new_vector.append(
                    current_vectors[j][k] + (
                    	temperature * random_vectors[j][k]))
            new_vectors.append(new_vector)
        normalize_vectors(new_vectors)

        # Accept change only if new_vectors produces a better epsilon
        epsilon = compute_epsilon(new_vectors)
        if epsilon < min_epsilon:
            min_epsilon = epsilon
            current_vectors = new_vectors
            i += 1

    return min_epsilon, current_vectors
\end{lstlisting}
\captionof{lstlisting}{The annealing algorithm used to generate optimum epsilon values}

To make this more precise, we present Python code for the algorithm above.

Even though this approach seems simple, it actually produces remarkably precise values -- this can be verified by cross-checking with results we've already proven / conjectured in this paper.

To make this more concrete, consider Table $1$.

The accuracy of our method can be noted by looking at the computed values for $n=2$ -- the computed values for different values of $N$ are very close to the expected value of $\cos^2 \pi/N$, that we already proved in Section $2$, which makes us confident about the accuracy of this numeric method.
\section{Applications to sphere packing}
We should be able to prove something about the radius of spheres you can pack in the case of $N=n+1$, taking as intuition square and hexagonal tilings of the plane and cubic and tetrahedral tilings of 3-space.

\begin{table}
   \centering
    \begin{tabular}{ | c | c | c | c |}
    \hline
      & $n=2$ & $n=3$ & $n=4$ \\ \hline
    $N=2$ & $0.0$ & $-$ & $-$ \\ \hline
    $N=3$ & $0.25$ & $0.0$ & $-$ \\ \hline
    $N=4$ & $0.5$ & $0.11$ & $0.0$ \\ \hline
    $N=5$ & $0.66$ & $0.21$ & $0.07$ \\ \hline
    $N=6$ & $0.75$ & $0.22$ & $0.14$ \\ \hline
    $N=7$ & $0.82$ & $0.36$ & $0.18$ \\ \hline
    $N=8$ &  $0.86$ & $0.44$ & $0.24$ \\ \hline
    $N=9$ &  $0.89$ & $0.49$ & $0.26$ \\ \hline
    $N=10$ & $0.91$ & $0.51$ & $0.32$ \\ \hline
    $N=11$ & $0.93$  & $0.55$ & $0.35$ \\ \hline
    $N=12$ & $0.94$ & $0.64$ & $0.41$ \\
    \hline
    \end{tabular}
    \caption {Minimum values of $\epsilon$ for different values of $(n, N)$, as computed by our annealing method}
\end{table}

\end{document}